Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-12-07 WholePlotTrajectories.Rmd
intro

Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and of the interactions among species such as competition [@Chesson2000].
Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Thereafter, models based on the observed trajectories after disturbance gave important insights into the recovery of ecosystem processes and services [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
It is recognized that determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape post-disturbance trajectories [@Carreno-Rocabado2012;@Poorter2016].
Community trajectories would then rely on deterministic processes and would be predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
Yet, observed trajectories in diversity and composition often deviate from predictions.
Besides, although the deterministic ecological processes at stake are identified, their relative importance compared to stochasticity remains debated [@Norden2015].
This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
Validations of the IDH remain scarce in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 
More ecological-meaningful analysis would couple richness with species evenness, which reveals the changes in the species abundance distribution, and (ii)species composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
Furthermore, a functional approach accounting for species biological attributes would directly link community diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
Eventually, a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines community resilience after disturbance [@Elmqvist2003; @Diaz2005].

>>>>

Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in the interactions among species [@Chesson2000].
Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
It is recognized that a determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape community trajectories afte disturbance [@Carreno-Rocabado2012;@Poorter2016].
Deterministic processes would then be major processes of community response to disturbance.
This would make post-disturbance trajectories deterministic, and then predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
Yet, observed trajectories in diversity and composition often deviate from predictions.
Although the deterministic ecological processes at stake are identified, They interplay with stochastic processes and their relative importance remains debated [@Norden2015].
This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
VThe IDH has been scarcely validated in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 
More ecological-meaningful analysis would couple richness with species evenness, which reveals the changes in the species abundance distribution, and (ii)species composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
Furthermore, a functional approach accounting for species biological attributes would directly link community diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
Eventually, a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines community resilience after disturbance [@Elmqvist2003; @Diaz2005].
